### 2020-12-17 submissions/genepi-cover/cover.tex
Adding cover letter (in progress)

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

I am submitting the manuscript ``XXX'' for your consideration as a Research Paper for publication in {\em Genetic Epidemiology}.

Penalized regression methods are an attractive approach for simultaneously selecting features and fitting a model based on those features, but have long been hampered by the difficulty of applying inferential tools.  In particular, it is of tremendous practical interest to be able to calculate a false discovery rate for the features that have been selected.  There have been several recent papers in this area proposing estimators of various kinds, but the focus has been almost entirely on least-squares regression; very little work has been done on estimating FDRs for penalized logistic regression or Cox regression models.  This article derives a general method for calculating the marginal false discovery rate for any likelihood-based regression model that satisfies the general regularity conditions for classical likelihood-based inference.  We have implemented this method in the R package {\tt ncvreg} and carefully investigate this approach for logistic and Cox regression via simulation and apply it to two high-dimensional studies, one involving binary outcomes and the other involving a time-to-event outcome.  I believe this manuscript would be of interest to the readers of {\em Biometrical Journal}, with the potential to become widely used in the field.

This work has not been previously published, is not under review elsewhere, and all authors concur with this submission.  Thank you very much for your consideration.

\EndLetter

\end{document}