### 2021-03-19 submissions/genepi-cover-1/cover.tex
Cover letter for resub

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

We are submitting the manuscript ``Penalized linear mixed models for structured genetic data'' for your consideration as a Review Article for publication in {\em Genetic Epidemiology}.

We feel that this is a review that the readers of {\em Genetic Epidemiology} will find important and timely.  Although one-SNP-a-time analyses are still the most common way of carrying out GWAS, genome-wide penalized regression methods such as the lasso are an increasingly common alternative.  However, little attention has been paid to correcting for population structure in such models.  This review covers this issue in detail, and describes the use of linear mixed models for penalized regression, a topic that has received little attention and been primarily confined to the plant and animal breeding literature.  Our review introduces and compares these methods, describing in detail how they work to reduce the potentially confounding influence of environmental factors in genetic association studies.

This work has not been previously published, is not under review elsewhere, and all authors concur with this submission.  Thank you very much for your consideration.

\EndLetter

\end{document}