### 2021-03-19 submissions/genepi-cover-2/cover.tex

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

We are resubmitting the manuscript ``Penalized linear mixed models for structured genetic data'' for your consideration as a Review Article for publication in {\em Genetic Epidemiology}.

Our revised manuscript addresses all of the issues raised by both reviewers.  In particular, we have added a section including a ``case study'' analysis applying the methods discussed in the paper to actual GWAS data to illustrate the performance of these method at a genome-wide scale.  Our point-by-point response to reviewer comments is included along with the revised manuscript.

Thank you very much for the opportunity to revise and resubmit our manuscript.  We feel that several improvements have been made to the original version and hope that you agree.

\EndLetter

\end{document}